The seasonal cycle of upper-ocean mixing at 8°N in the Bay of Bengal

### 2019-11-19 paper1.tex
Final edits to text.

\abstract{
We describe the seasonal cycle of mixing in the top \SIrange{30}{100}{m} of the Bay of Bengal as observed by moored mixing meters (χpods) deployed along 8°N between 85.5°E and 88.5°E in 2014 and 2015.
All χpod observations were combined to form seasonal-mean vertical profiles of turbulence diffusivity $K_T$ in the top 100m.
The strongest turbulence is observed between July and November during the southwest and post-monsoon seasons.
The northeast monsoon (December -- February) is a period of similarly high mean $K_T$ but an order of magnitude lower median $K_T$, a sign of energetic episodic mixing events forced by near-inertial shear events.
The months of March and April, a period of weak wind forcing and low near-inertial shear amplitude, are characterized by near-molecular values of $K_T$ in the thermocline for weeks at a time.
High mixing events coincide with the passage of surface-forced downward-propagating near-inertial waves and with the presence of enhanced low-frequency shear associated with the Summer Monsoon Current and other mesoscale features between July and October.
This seasonal cycle of mixing is consequential. 
We find that monthly-averaged turbulent transport of salt out of the salty Arabian Sea water between August and January is significant relative to local $E-P$.
The magnitude of this salt flux is approximately that required to close model-based salt budgets for the upper Bay of Bengal.
}
\maketitle

@@

\begin{comment}
\label{sec:org2d483eb}

\begin{itemize}
\item \textbf{reviewer 1didn't like previous introduction}. comments from reviewer 1 below:
\item "there is precious little rationale for the study provided, nor is the context of previous work much discussed in the introduction.  In fact, after a cursory first paragraph introducing the Bay of Bengal, its rivers and model predictions that vertical mixing during the Southwest monsoon is important for maintaining the Bay's salt budget, the study pretty launches into a presentation of its first results. These are the calculations of wind stress, altimetric eke,  mean surface currents, inertial energy flux and halocline dept that are intended to provide context for the study.  The problem with all of this is two-fold. Firstly, unless the authors only intend to reach readers interested in the Bay of Bengal and perhaps some upper ocean mixing people, then their introduction needs to be much broader.  Why should/would anyone else care what happens in this small region of the Indian Ocean?  Does this have a broader impact on livelihoods, ecosystems and climate change?"
\item "Secondly, in using their first results to provide context for the study, the authors commit a cardinal sin of failing to adequately describe their methods so their results can be reproduced.  The authors assume their readers are all familiar details of the datasets and non-trivial methods used - AVISO, OSCAR, MERRA-2, Alford slab-ocean mixed layer model, MIMOC etc.  These maybe widely-used datasets, but you can't assume everyone knows what they are, or what biases you may have introduced by choices in averaging, or drag coefficients. Also a few previous studies have been cited here, but as many of them would probably have done better more detailed jobs of explaining the circulation, stratification and monsoons of the Bay of Bengal, perhaps their original results should have been discussed instead of the perfunctory presentation of these we are gifted with here."
\item Objective:  This manuscript would benefit from a clear statement of the objective of the study.  Is it to provide a description of the seasonal cycle in upper ocean mixing here or address the question of whether SW monsoon-driving mixing can provide Arabian sea salt fluxes to close the  surface layer salt budgets?  Which is it? And if it is the former, then as an overarching objective, how was this justified by your rationale?
\end{itemize}

\end{comment}

The Bay of Bengal (the Bay) is the eastern semi-enclosed basin of the north Indian Ocean. 
The shallow salinity-controlled stratification in the upper Bay allows for rapid coupling with the atmosphere, and modulation of sea surface temperature (SST) within the Bay of Bengal has been linked to variations in the South Asian monsoon \citep[e.g.,][]{Vecchi2002,Roxy2014}. 
The influence of processes controlling upper ocean stratification thus extend beyond the physical footprint of the Bay. 
The Bay has a particularly strong influence on rainy and dry periods , over the Indian subcontinent, termed active and break periods respectively --- much of central India's annual rainfall results from convective systems that originate over the Bay and then propagate northwestward over the Indian subcontinent between June and September \citep{Gadgil2003a,Goswami2003}.
Interannual variations in mean rainfall are strongly correlated with fluctuations in India's agricultural output \citep{Gadgil2006}, lending significant social relevance to the problem of understanding air-sea interaction and near-surface ocean dynamics that influence the Bay's SST.

@@

The Indian Ocean monsoon and its associated precipitation is visualized in Figure \ref{fig:ppt} using seasonal mean wind stress from the Tropflux estimate \citep{Kumar2012} and precipitation from the Tropical Rainfall Measuring Mission (TRMM) Multi-satellite Precipitation Analysis dataset \citep{trmm}.
Between May and September (southwest or SW monsoon), the winds are strong and southwesterly throughout the Indian Ocean basin.
Precipitation over the Indian subcontinent is substantial (Figure \ref{fig:ppt}c).
The months of October and November (post-monsoon period, SWNE) are characterized by weak mean wind stress over most of the basin including the Bay (Figure \ref{fig:ppt}d). 
The seasonal averaging hides the episodic presence of strong cyclones in the Bay that bring large amounts of rain and significantly affect lives of those residing along the perimeter of the Bay.
Recent examples of cyclones that strengthened over the Bay and made landfall resulting in loss of life and severe damages, include Category 5 Tropical Cyclone Phailin in October 2013 and Category 4 Tropical Cyclone Hudhud in October 2014.
% Sri Lanka in particular receives a substantial fraction of its annual mean precipitation during this season.
The mean winds strengthen in December and switch to being northeasterly up until February (northeast or NE monsoon).
These winds are weaker than those during the SW monsoon.
Nations on the Bay's Rim i.e. India and Sri Lanka receive relatively little rainfall during this period and the precipitation maximum is located over the Bay (Figure \ref{fig:ppt}a).
The months of March and April are a period of weak winds and almost no precipitation north of 4°N (northeast-southwest transition, NESW; Figure \ref{fig:ppt}b).

The monsoon cycle of wind forcing described above imprints seasonality on the Bay's circulation (\citealp{Schott2002,Shankar2002}).
The East India Coastal Current (EICC) spins up at the Bay's western boundary during both monsoons, flowing northward between May and October and then southward between December and April.
The EICC is readily visible in seasonally-averaged estimates of near-surface ocean velocity (vectors in Figure \ref{fig:spatial}a--e from the Ocean Surface Current Analysis Real-time product, OSCAR; \citealp{Bonjean2002}).
(OSCAR is a diagnostic estimate at five-day frequency that ignores local acceleration and nonlinearities but accounts for geostrophic, thermal wind and Ekman currents.)
The EICC exists as a discontinuous flow with many recirculation loops and is visible as a local maximum along India's eastern coast in maps of geostrophic eddy kinetic energy \(\text{EKE} = 0.5[u_g^2 + v_g^2]\) (colored field in Figure \ref{fig:spatial}a--e; \citealp{Durand2009}).
Here \((u_g, v_g)\) are geostrophic velocity anomalies computed from delayed-time sea surface height estimates as measured by multiple satellite altimeters by the Copernicus Marine Environment Monitoring Service (CMEMS)\footnote{DT all-sat-merged Global Ocean Gridded SSALTO/DUACS Sea Surface Height L4 product and derived variables (dataset-duacs-rep-global-merged-allsat-phy-l4-v3).}. 
Apart from the EICC, one other major circulation feature is the Summer or Southwest Monsoon Current (SMC).
The SMC is visible in the seasonal mean during the SW monsoon as an eastward jet along 8°N between 85°E and 92°E in Figure \ref{fig:spatial}d (vectors).
Peak velocity in the SMC can exceed \SI{1.5}{\meter\per\second} and northward transport has been estimated to be in the range \SIrange{10}{27}{Sv} though some of this transport may be due to recirculations \citep{Wijesekera2016a,Vinayachandran1999,Webber2018}.
The southwestern and south-central Bay is a site of energetic mesoscale variability during the SW monsoon (our Figure \ref{fig:spatial}d; \citealp{Chen2018}).
The elevated EKE reflects at least three mesoscale features: a westward propagating anticyclone \citep{Wijesekera2016a}, a large cyclonic eddy that spins up annually off the coast of Sri Lanka (the Sri Lanka Dome, \citealp{Vinayachandran1998}) and the SMC which threads a pathway between the Sri Lanka Dome to the north and the anticyclone to the south \citep{Vinayachandran1999,Wijesekera2016a}.
During the NE monsoon, the mean circulation in southern Bay reverses and the Northeast Monsoon Current (NMC) flows westward with a weaker signal in EKE (Figure \ref{fig:spatial}a,b).

@@

All presented turbulence quantities were obtained using χpods: self-contained instruments each consisting of two fast-response FP-07 thermistors, a pitot-static tube for high-frequency speed measurements, a pressure sensor, a compass, and accelerometers \citep{Moum2009a,Moum2015}.
Refinement over many years has resulted in a system that can return records of turbulent temperature fluctuations for up to a year or more.
The two thermistors on the χpod record temperature fluctuations at 100Hz.
Temperature gradient spectra are computed using 1 second data intervals and are fit to the theoretical spectrum of \cite{Kraichnan1968} in the viscous-convective range \citep{Moum2009a}.
The Kraichnan spectrum is a function of two quantities: the turbulence dissipation rate of temperature variance \(χ\) and the turbulence dissipation rate of kinetic energy \(ε\), but the χpods only record one quantity, temperature.
The dependence on \(ε\) arises from the \cite{Batchelor1959} wavenumber which marks the end of the viscous-convective range.
Since χpod thermistors do not resolve the Batchelor wavenumber typically (e.g. \citealp{Lueck1977}), fitting the Kraichnan spectrum requires specification of \(ε\).
In the absence of an independent estimate of \(ε\), we assume that the turbulence diffusivities of temperature  \(K^t_T = χ/2/T_z^2\) and density \(K^t_ρ = Γε/N^2\) are equal with mixing efficiency \(Γ = 0.2\) for stratified turbulence (\citealp{Osborn1972}; \citealp{Osborn1980}; and \citealp{Gregg2018} respectively).
This yields a relationship between \(χ\) and \(ε\), 
\begin{equation}
    \label{eq:eps}
    ε = \frac{N² χ}{2 Γ T_z²}
\end{equation}
and a solution is obtained by fitting the spectrum through the iterative procedure described in \cite{Moum2009a}.
The buoyancy frequency \(N\) and vertical temperature gradient \(T_z\) are estimated using two CTD instruments deployed above and below the χpod.
\emph{In-situ} comparisons between χpod estimates and more ``standard'' estimates from vertical microstructure profiles are favorable under stably-stratified sheared conditions \citep{Perlin2012,Pujiana2018}.
Total temperature and salt diffusivities \(K_T, K_S\), heat flux \(J_q^t\) and salt flux \(J_s^t\) are estimated from a time series of \(χ\) using
\begin{subequations}
\label{eq:def}
 \begin{align}
  K_T &= κ_T(S,T,P) + \frac{χ/2}{T_z^2}, \\
  K_s &= κ_s + \frac{χ/2}{T_z^2}, \\
  J_q^t &= - ρ_0\, c_p\, K_T\, T_z, \\
  J_s^t &= - ρ_0 \, K_S \, S_z;
\end{align}
\end{subequations}
where \(κ_T, κ_s\) are the molecular diffusivity of temperature and salinity respectively, and \(T_z, S_z\) are background temperature and salinity gradients (usually obtained by differencing nearby CTDs on the moorings; subscript \(z\) indicates \(z\)-derivative).
Again we have assumed that high Reynolds number geophysical turbulence mixes all scalars at the same rate so that the \emph{turbulence} diffusivities of both temperature and salinity are equal i.e. \((χ/2)/T_z^2\).

A challenge with analyzing χpods deployed in the Bay's thermocline is the frequent occurrence of weakly turbulent and near-laminar flow for extended periods of time as has been recorded with microstructure measurements in the Aegean Sea \citep{Gregg2012} and in the Arctic \citep{Scheifele2018}.
Analyzing microstructure measurements in such environments is challenging given that the usual assumptions of isotropy, steadiness and homogeneity break down \citep{Rohr1988,Itsweire1993,Gargett1984a}.
The χpod records ``bit noise'' when the turbulent temperature fluctuations are below the FP-07 sensor's detection threshold.
We can account for such behaviour using knowledge of the  circuit components involved (Appendix B).
When the recorded temperature variance of an one second subset of data is within an arbitrary factor of 1.5 of the inferred noise variance of the sensor, we set \(ε\) to NaN and \(χ\) to 0 resulting in \(K_T, K_S\)  being set to molecular values \(κ_T, κ_S\) and the resulting fluxes \(J_q^t, J_s^t\) being that due to molecular diffusion (eqn. \ref{eq:def}).
We do so following \cite{Gregg2012} with the understanding that setting \(χ\) to any non-zero value during such periods seems unjustifiable.

@@

Two Teledyne RD Instruments ADCPs were deployed at the top of each mooring: an upward-looking Workhorse \SI{300}{kHz} sampling every half hour in \SI{2}{m} bins and a downward-looking Long Ranger \SI{75}{kHz} sampling every hour in \SI{8}{m} bins (further details are available in \citealp{Wijesekera2016a}).
A data gap in velocity coverage exists between the two ADCPs that is approximately \SI{21}{m} wide.
The shallower χpod was deployed within the blanking zone of the downward looking ADCP, so shear can be directly estimated only at the deeper χpod.
We estimate shear by first interpolating the velocities over the gap in depth, central differencing the interpolated velocity over three \SI{8}{m} wide bins, and then reintroducing the gap.
Each mooring contained more than fifteen temperature sensors of various kinds distributed between the buoy and \SI{352}{m} below the buoy.
Salinity coverage was coarser with four sensors deployed within a \SI{50}{m} depth below the buoy and one sensor at \SI{352}{m} \citep{Wijesekera2016a}.
Three of the four shallow salinity sensors were concentrated around the two χpods deployed \SI{12}{m} and \SI{32}{m} below the buoy.

@@

Much of this manuscript describes a seasonal cycle in thermocline turbulence that coincides with a seasonal cycle in thermocline shear.
The seasonal variation in turbulence will be discussed along with the seasonal variation in the shear field, decomposed into three components as described below.
Bursts in near-inertial shear will be linked back to surface winds using an approximate estimate of mixed-layer wind energy input obtained using a slab mixed layer model, also described below.
First we introduce and rationalize our decomposition of the shear field.
%Then we describe our construction of a seasonally averaged vertical profile (Section \ref{sec:results}\ref{sec:profile}).
%Later in Section \ref{sec:results}\ref{sec:nrl}, we present time series of turbulence and shear at a single χpod and discuss the seasonal cycle implied by each of these figures in more detail (Section \ref{sec:results}\ref{sec:seasonal}).

@@

At all three moorings, \emph{Eulerian} rotary spectra of vertical shear \(S\tot = \sqrt{u_z^2 + v_z^2}\) at \SI{152}{m} depth\footnote{We choose \SI{152}{m} to avoid any uncertainties with having to interpolate over the gap in ADCP coverage} are dominated by a broad peak at \(-f_0\) (40\%--60\% of sampled variance), narrow secondary peaks at \(-f_0 \pm ω_{M2}\) (\(ω_{M2}\) is the \(M_2\) tidal frequency, 5\%--10\% variance) and distributed variance at frequencies less than \SI{10}{days} reflecting meanders of the Summer Monsoon Current (20\% variance).
These spectra are presented in Figure \ref{fig:nrlspectra}a,c,e (clockwise in black, counterclockwise in red).
The narrow peaks at \(-f_0 \pm ω_{M2}\) are a sign of vertical advection or pumping of near-inertial shear layers by the \(M_2\) tide which Doppler-shifts spectral energy from \(-f_0\) to \(-f_0 \pm ω_{M2}\) \citep{Alford2001a}.
The effect of tidal pumping can be removed by estimating the spectra in isopycnal space (e.g. \citealp{Alford2017}).
Given the sparse sampling in salinity, we instead estimate spectra in \emph{isothermal} space.
The peaks at \(f_0 \pm ω_{M2}\) are much less prominent at the $T=\SI{18}{\celsius}$ isotherm at all moorings (annual mean depth 150m; Figure \ref{fig:nrlspectra}b,d,f), leading us to interpret the near-tidal peaks in the Eulerian spectra (Figure \ref{fig:nrlspectra}a,c,e) as primarily being near-inertial shear that is Doppler shifted to near-tidal frequencies.
Ideally we would interpret the χpod mixing estimates using a time series of isothermal shear that is filtered to isolate the low frequency and near-inertial components.
It is not possible to obtain a gapless estimate of these filtered components given the gap in ADCP coverage.
Instead we proceed by conducting our analysis in the Eulerian frame as follows.

@@

Depth-time maps of the mean squared shear for three shear components \(S\low, S\niwp, S\res\) along with the total shear \(S\tot\) are shown in Figure \ref{fig:shears}.
At all three moorings, energetic shear is observed in January, February, and for an extended period between July and November.
The shear field is relatively weak between mid-March and the beginning of June.
Episodic energetic bursts in near-inertial shear are seen at all three moorings outside these months.
All three moorings see a large rise in low-frequency shear between July and November.
This is an indication of the Sri Lanka Dome and a large anticyclonic eddy through the array as the meandering of the Summer Monsoon Current (note EKE maximum inferred from altimetric data in Figure \ref{fig:spatial}a--e).
The magnitude of the low-frequency shear is comparable to that of near-inertial shear at all three locations during the SW monsoon.
% At NRL4, the low-frequency, near-inertial and residual components contribute equally to total shear.
The residual \(S\res\) is generally weak relative to the \(S\low\) and \(S\niwp\).
The episodic nature of near-inertial shear events prevent a confident estimation of the magnitude of its seasonal cycle given that only one annual cycle is resolved.
However the seasonal signal in \emph{total} shear — high shear between June and February, and low shear between March and June — is robust and consistent across all three moorings.

@@

Despite these caveats, Figure \ref{fig:vert} presents a useful summary of observed mixing along 8°N.
There is a clear seasonal cycle in turbulent diffusivity in the upper 30--100m at the mooring locations just as at NRL5 (Figure \ref{fig:nrl}).
Vertical profiles of both mean and median values of \(K_T\) are always surface intensified (tables of both means and medians are provided in Appendix B).
The amplitude of the seasonal cycle in mean diffusivities is roughly an order of magnitude with mean $K_T \approx \SI{1e-4}{\metre\squared\per\second}$ during both monsoons.
Median $K_T$ is approximately an order of magnitude larger during the SW monsoon as compared to the NE monsoon (\SI{1e-6}{\metre\squared\per\second} versus \SI{1e-7}{\metre\squared\per\second}) indicating that energetic mixing events are rarer during the NE monsoon.
The most striking feature of Figure \ref{fig:vert} is the near-complete lack of mixing in the south-central Bay's thermocline during the months of March and April --- median diffusivity values are only slightly greater than molecular diffusivity \(κ_T\) at depths greater than \SI{60}{m}.
The observation of near-molecular diffusivity at the deep χpod at NRL5 is thus consistent across the other two moorings.

>>>>

\abstract{
We describe the seasonal cycle of mixing in the top \SIrange{30}{100}{m} of the Bay of Bengal as observed by moored mixing meters (χpods) deployed along 8°N between 85.5°E and 88.5°E in 2014 and 2015.
All χpod observations were combined to form seasonal-mean vertical profiles of turbulence diffusivity $K_T$ in the top 100m.
The strongest turbulence is observed during the southwest and post-monsoon seasons ie. between July and November.
The northeast monsoon (December -- February) is a period of similarly high mean $K_T$ but an order of magnitude lower median $K_T$, a sign of energetic episodic mixing events forced by near-inertial shear events.
The months of March and April, a period of weak wind forcing and low near-inertial shear amplitude, are characterized by near-molecular values of $K_T$ in the thermocline for weeks at a time.
Strong mixing events coincide with the passage of surface-forced downward-propagating near-inertial waves and with the presence of enhanced low-frequency shear associated with the Summer Monsoon Current and other mesoscale features between July and October.
This seasonal cycle of mixing is consequential.
We find that monthly-averaged turbulent transport of salt out of the salty Arabian Sea water between August and January is significant relative to local $E-P$.
The magnitude of this salt flux is approximately that required to close model-based salt budgets for the upper Bay of Bengal.
}
\maketitle

@@

The Bay of Bengal (the Bay) is the eastern semi-enclosed basin of the north Indian Ocean. 
The shallow salinity-controlled stratification in the upper Bay allows for rapid coupling with the atmosphere, and modulation of sea surface temperature (SST) within the Bay of Bengal has been linked to variations in the South Asian monsoon \citep[e.g.,][]{Vecchi2002,Roxy2014}. 
The influence of processes controlling upper ocean stratification thus extend beyond the physical footprint of the Bay. 
The Bay has a particularly strong influence on rainy and dry periods over the Indian subcontinent, termed active and break periods respectively.
Much of central India's annual rainfall results from convective systems that originate over the Bay and then propagate northwestward over the Indian subcontinent between June and September \citep{Gadgil2003a,Goswami2003}.
Interannual variations in mean rainfall are strongly correlated with fluctuations in India's agricultural output \citep{Gadgil2006}, lending significant social relevance to the problem of understanding air-sea interaction and near-surface ocean dynamics that influence the Bay's SST.

@@

The Indian Ocean monsoon and its associated precipitation is visualized in Figure \ref{fig:ppt} using seasonal mean wind stress from the Tropflux estimate \citep{Kumar2012} and precipitation from the Tropical Rainfall Measuring Mission (TRMM) Multi-satellite Precipitation Analysis dataset \citep{trmm}.
Between May and September (southwest or SW monsoon), the winds are strong and southwesterly throughout the Indian Ocean basin.
Precipitation over the Indian subcontinent is substantial (Figure \ref{fig:ppt}c).
The months of October and November (post-monsoon period, SWNE) are characterized by weak mean wind stress over most of the basin including the Bay (Figure \ref{fig:ppt}d). 
Seasonal averaging hides the episodic presence of strong cyclones in the Bay that bring large amounts of rain and significantly affect lives of those residing along the perimeter of the Bay.
Recent examples of cyclones that strengthened over the Bay and made landfall resulting in loss of life and severe damages, include Category 5 Tropical Cyclone Phailin in October 2013 and Category 4 Tropical Cyclone Hudhud in October 2014.
% Sri Lanka in particular receives a substantial fraction of its annual mean precipitation during this season.
The mean winds strengthen in December and switch to being northeasterly up until February (northeast or NE monsoon).
These winds are weaker than those during the SW monsoon.
Nations on the Bay's Rim i.e. India and Sri Lanka receive relatively little rainfall during this period and the precipitation maximum is located over the Bay (Figure \ref{fig:ppt}a).
The months of March and April are a period of weak winds and almost no precipitation north of 4°N (northeast-southwest transition, NESW; Figure \ref{fig:ppt}b).

The monsoon imprints seasonality on the Bay's circulation (\citealp{Schott2002,Shankar2002}).
The East India Coastal Current (EICC) spins up at the Bay's western boundary during both monsoons, flowing northward between May and October and then southward between December and April.
The EICC is readily visible in seasonally-averaged estimates of near-surface ocean velocity (vectors in Figure \ref{fig:spatial}a--e from the Ocean Surface Current Analysis Real-time product, OSCAR; \citealp{Bonjean2002})\footnote{OSCAR is a diagnostic estimate of near-surface velocity at five-day frequency that ignores local acceleration and nonlinearities but accounts for geostrophic, thermal wind and Ekman currents.}
The EICC exists as a discontinuous flow with many recirculation loops and is visible as a local maximum along India's eastern coast in maps of geostrophic eddy kinetic energy \(\text{EKE} = 0.5[u_g^2 + v_g^2]\) (colored field in Figure \ref{fig:spatial}a--e; \citealp{Durand2009}).
Here \((u_g, v_g)\) are geostrophic velocity anomalies computed from delayed-time sea surface height estimates as measured by multiple satellite altimeters by the Copernicus Marine Environment Monitoring Service (CMEMS)\footnote{DT all-sat-merged Global Ocean Gridded SSALTO/DUACS Sea Surface Height L4 product and derived variables (dataset-duacs-rep-global-merged-allsat-phy-l4-v3).}. 
Apart from the EICC, one other major circulation feature is the Summer or Southwest Monsoon Current (SMC).
The SMC is visible in the seasonal mean during the SW monsoon as an eastward jet along 8°N between 85°E and 92°E in Figure \ref{fig:spatial}d (vectors).
Peak velocity in the SMC can exceed \SI{1.5}{\meter\per\second} and northward transport has been estimated to be in the range \SIrange{10}{27}{Sv}, likely an overestimate due to the presence of recirculations \citep{Wijesekera2016a,Vinayachandran1999,Webber2018}.
The southwestern and south-central Bay is a site of energetic mesoscale variability during the SW monsoon (our Figure \ref{fig:spatial}d; \citealp{Chen2018}).
The elevated EKE reflects at least three mesoscale features: a westward propagating anticyclone \citep{Wijesekera2016a}, a large cyclonic eddy that spins up annually off the coast of Sri Lanka (the Sri Lanka Dome, \citealp{Vinayachandran1998}) and the SMC which threads a pathway between the Sri Lanka Dome to the north and the anticyclone to the south \citep{Vinayachandran1999,Wijesekera2016a}.
During the NE monsoon, the mean circulation in southern Bay reverses and the Northeast Monsoon Current flows westward with a weaker signal in EKE (Figure \ref{fig:spatial}a,b).

@@

All presented turbulence quantities were obtained using χpods: self-contained instruments each consisting of two fast-response FP-07 thermistors, a pitot-static tube for high-frequency speed measurements, a pressure sensor, a compass, and accelerometers \citep{Moum2009a,Moum2015}.
Refinement over many years has resulted in a system that can return records of turbulent temperature fluctuations for up to a year or more.
The two thermistors on the χpod record temperature fluctuations at 100Hz.
Temperature gradient spectra are computed using 1 second data intervals and are fit to the theoretical spectrum of \cite{Kraichnan1968} in the viscous-convective range \citep{Moum2009a}.
The Kraichnan spectrum is a function of two quantities: the turbulence dissipation rate of temperature variance \(χ\) and the turbulence dissipation rate of kinetic energy \(ε\), but the χpods only record one quantity, temperature.
The dependence on \(ε\) arises from the \cite{Batchelor1959} wavenumber \( (ε/νk_T²)^{1/4} \) which marks the end of the viscous-convective range.
Since χpod thermistors do not resolve the Batchelor wavenumber typically (e.g. \citealp{Lueck1977}), fitting the Kraichnan spectrum requires specification of \(ε\).
In the absence of an independent estimate of \(ε\), we assume that the turbulence diffusivities of temperature  \(K^t_T = χ/2/T_z^2\) and density \(K^t_ρ = Γε/N^2\) are equal with mixing efficiency \(Γ = 0.2\) for stratified turbulence (\citealp{Osborn1972}; \citealp{Osborn1980}; and \citealp{Gregg2018} respectively).
This yields a relationship between \(χ\) and \(ε\), 
\begin{equation}
    \label{eq:eps}
    ε = \frac{N² χ}{2 Γ T_z²}
\end{equation}
and a solution is obtained by fitting the spectrum through the iterative procedure described in \cite{Moum2009a}.
The buoyancy frequency \(N\) and vertical temperature gradient \(T_z\) are estimated using two CTD instruments deployed above and below the χpod.
\emph{In-situ} comparisons between χpod estimates and more ``standard'' estimates from vertical microstructure profiles are favorable under stably-stratified sheared conditions \citep{Perlin2012,Pujiana2018}.
Total temperature and salt diffusivities \(K_T, K_S\), heat flux \(J_q^t\) and salt flux \(J_s^t\) are estimated from a time series of \(χ\) using
\begin{subequations}
\label{eq:def}
 \begin{align}
  K_T &= κ_T(S,T,P) + \frac{χ/2}{T_z^2}, \\
  K_s &= κ_s + \frac{χ/2}{T_z^2}, \\
  J_q^t &= - ρ_0\, c_p\, K_T\, T_z, \\
  J_s^t &= - ρ_0 \, K_S \, S_z;
\end{align}
\end{subequations}
where \(κ_T, κ_s\) are the molecular diffusivity of temperature and salinity respectively, and \(T_z, S_z\) are background temperature and salinity gradients (usually obtained by differencing nearby CTDs on the moorings; subscript \(z\) indicates \(z\)-derivative).
Again we have assumed that high Reynolds number geophysical turbulence mixes all scalars at the same rate so that the \emph{turbulence} diffusivities of both temperature and salinity are equal i.e. \((χ/2)/T_z^2\).

A challenge with analyzing χpods deployed in the Bay's thermocline is the frequent occurrence of weakly turbulent and near-laminar flow for extended periods of time as has been recorded with microstructure measurements in the Aegean Sea \citep{Gregg2012} and in the Arctic \citep{Scheifele2018}.
Analyzing microstructure measurements in such environments is challenging given that the usual assumptions of isotropy, steadiness and homogeneity break down \citep{Rohr1988,Itsweire1993,Gargett1984a}.
In weakly turbulent environments, the χpod records ``bit noise'' when the turbulent temperature fluctuations are below the FP-07 sensor's detection threshold.
We can account for such behaviour using knowledge of the  circuit components involved (Appendix B).
When the recorded temperature variance of an one second subset of data is within an arbitrary factor of 1.5 of the inferred noise variance of the sensor, we set \(ε\) to NaN and \(χ\) to 0 resulting in total diffusivities \(K_T, K_S\)  being set to molecular values \(κ_T, κ_S\) and the resulting fluxes \(J_q^t, J_s^t\) being that due to molecular diffusion (eqn. \ref{eq:def}).
We do so following \cite{Gregg2012} with the understanding that setting \(χ\) to any non-zero value during such periods seems unjustifiable.

@@

Two Teledyne RD Instruments ADCPs were deployed at the top of each mooring: an upward-looking Workhorse \SI{300}{kHz} sampling every half hour in \SI{2}{m} bins and a downward-looking Long Ranger \SI{75}{kHz} sampling every hour in \SI{8}{m} bins (further details are available in \citealp{Wijesekera2016a}).
A data gap in velocity coverage exists between the two ADCPs that is approximately \SI{21}{m} wide.
The shallower χpod was deployed within the blanking zone of the downward looking ADCP, so shear can be directly estimated only at the deeper χpod.
We estimate shear by first linearly interpolating the velocities over the gap in depth, central differencing the interpolated velocity over three \SI{8}{m} wide bins, and then reintroducing the gap.
Each mooring contained more than fifteen temperature sensors of various kinds distributed between the buoy and \SI{352}{m} below the buoy.
Salinity coverage was coarser with four sensors deployed within a \SI{50}{m} depth below the buoy and one sensor at \SI{352}{m} \citep{Wijesekera2016a}.
Three of the four shallow salinity sensors were concentrated around the two χpods which were deployed \SI{12}{m} and \SI{32}{m} below the buoy.

@@

We now describe a seasonal cycle in thermocline turbulence that coincides with a seasonal cycle in thermocline shear.
The seasonal variation in turbulence will be discussed along with the seasonal variation in the shear field, decomposed into three components as described below.
Bursts in near-inertial shear will be linked back to surface winds using an approximate estimate of mixed-layer wind energy input obtained using a slab mixed layer model, also described below.
First we introduce and rationalize our decomposition of the shear field.
%Then we describe our construction of a seasonally averaged vertical profile (Section \ref{sec:results}\ref{sec:profile}).
%Later in Section \ref{sec:results}\ref{sec:nrl}, we present time series of turbulence and shear at a single χpod and discuss the seasonal cycle implied by each of these figures in more detail (Section \ref{sec:results}\ref{sec:seasonal}).

@@

At all three moorings, \emph{Eulerian} rotary spectra of vertical shear \(S\tot = \sqrt{u_z^2 + v_z^2}\) at \SI{152}{m} depth\footnote{We choose \SI{152}{m} to avoid any uncertainties associated with interpolating over the gap in ADCP coverage.} are dominated by a broad peak at \(-f_0\) (40\%--60\% of sampled variance), narrow secondary peaks at \(-f_0 \pm ω_{M2}\) (\(ω_{M2}\) is the \(M_2\) tidal frequency, 5\%--10\% variance) and distributed variance at frequencies less than \SI{10}{days} reflecting meanders of the Summer Monsoon Current (20\% variance).
These spectra are presented in Figure \ref{fig:nrlspectra}a,c,e (clockwise in black, counterclockwise in red).
The narrow peaks at \(-f_0 \pm ω_{M2}\) are a sign of vertical advection or pumping of near-inertial shear layers by the \(M_2\) tide which Doppler-shifts spectral energy from \(-f_0\) to \(-f_0 \pm ω_{M2}\) \citep{Alford2001a}.
The effect of tidal pumping can be removed by estimating the spectra in isopycnal space (e.g. \citealp{Alford2017}).
Given the sparse sampling in salinity, we instead estimate spectra in \emph{isothermal} space.
The peaks at \(f_0 \pm ω_{M2}\) are much less prominent at the $T=\SI{18}{\celsius}$ isotherm at all moorings (annual mean depth 150m; Figure \ref{fig:nrlspectra}b,d,f), leading us to interpret the near-tidal peaks in the Eulerian spectra (Figure \ref{fig:nrlspectra}a,c,e) as primarily being near-inertial shear that is Doppler shifted to near-tidal frequencies.
Ideally we would interpret the χpod mixing estimates using a time series of isothermal shear that is filtered to isolate the low frequency and near-inertial components.
It is not possible to obtain a gapless estimate of these filtered components given the gap in ADCP coverage.
Instead we proceed by conducting our analysis in the Eulerian frame as follows.

@@

Depth-time maps of the mean squared shear for three shear components \(S\low, S\niwp, S\res\) along with the total shear \(S\tot\) are shown in Figure \ref{fig:shears}.
At all three moorings, energetic shear is observed in January, February, and for an extended period between July and November.
The shear field is relatively weak between mid-March and the beginning of June.
Episodic energetic bursts in near-inertial shear are seen at all three moorings outside these months.
All three moorings see a large rise in low-frequency shear between July and November.
This is an indication of the Sri Lanka Dome and a large anticyclonic eddy through the array as the meandering of the Summer Monsoon Current (note EKE maximum inferred from altimetric data in Figure \ref{fig:spatial}a--e).
The magnitude of the low-frequency shear is comparable to that of near-inertial shear at all three locations during the SW monsoon.
% At NRL4, the low-frequency, near-inertial and residual components contribute equally to total shear.
The residual \(S\res\) is generally weak relative to the \(S\low\) and \(S\niwp\).
The episodic nature of near-inertial shear events prevent a confident estimation of the magnitude of its seasonal cycle given that only one full annual cycle was recorded.
However the seasonal signal in \emph{total} shear — high shear between June and February, and low shear between March and June — is robust and consistent across all three moorings.

@@

Despite these caveats, Figure \ref{fig:vert} presents a useful summary of observed mixing along 8°N.
There is a clear seasonal cycle in turbulent diffusivity in the upper 30--100m at all mooring locations that mirrors the seasonal cycle at NRL5 in Figure \ref{fig:nrl}.
Vertical profiles of both mean and median values of \(K_T\) are always surface intensified (tables of both means and medians are provided in Appendix B).
The amplitude of the seasonal cycle in mean diffusivities is roughly an order of magnitude with mean $K_T \approx \SI{1e-4}{\metre\squared\per\second}$ during both monsoons.
Median $K_T$ is approximately an order of magnitude larger during the SW monsoon as compared to the NE monsoon (\SI{1e-6}{\metre\squared\per\second} versus \SI{1e-7}{\metre\squared\per\second}) indicating that energetic mixing events are rarer during the NE monsoon.
The most striking feature of Figure \ref{fig:vert} is the near-complete lack of mixing in the south-central Bay's thermocline during the months of March and April --- median diffusivity values are only slightly greater than molecular diffusivity \(κ_T\) at depths greater than \SI{60}{m}.
The observation of near-molecular diffusivity at the deep χpod at NRL5 is thus consistent across the other two moorings.
